Transfer Learning in Visual and Relational Reasoning

### 2020-02-08 Transfer_Learning/cvpr_2nd_submission/experiments_reasoning_transfer.tex

\subsection{Reasoning transfer in CLEVR-CoGenT}
\label{sec:reasoning-transfer-clevr}
In the following we have relied on the CoGent-A variant only.
Using the question groups (categories) defined by the authors, we reorganized training and testing splits with the goal of measuring whether mastering reasoning for one group can help with learning of the other.

@@

Next, we trained and tested 5 SAMNet instances, each on a single task group $t$.
Interestingly, we observe significant accuracy drops (from 18 up to almost 35 points) in all task groups aside of the \textit{Query Attribute}.
We hypotesize this results from the fact that mastering visual grounding and recognizing object attributes (being the core of the latter task) is a necessary skill the system needed by the other tasks.
Thus training on \textit{Query Attribute} helps others, whereas without any training on that particular group the system struggles with developing such a skill.

@@

\subsection{Reasoning transfer in COG Canonical}
\label{sec:reasoning-transfer-cog}

We conduct the following experiments using the Canonical variant of COG to study whether reasoning transfer learning was effective in leveraging information gained by training a task family at a higher level of the hierarchy.
Order of the performed experiments is analogical to the one from previous chapter.
First column (\cref{fig:cog_reasoning_transfer}, "Train on All") is our baseline, i.e.
we train a single SAMNet instance on all tasks and test on each of the five task groups from the task hierarchy presented in \cref{fig:task-groups}.
Please note that those results are in fact weighted averages when aggregating accuracies of our model on the Canonical variant (\cref{fig:samnet_cog_detailed}).

Next

 \begin{itemize}
 	\compresslist
	\item Train and test SAMNet on each of the 5 task groups at the lowest level of the hierarchy (Single-task training);
	\item Jointly train on:
	\begin{itemize}
		\item \textbf{Group A} and test on each task from the lowest groups (i.e. \textbf{Basic}, \textbf{Obj-Attr} and \textbf{Compare}) separately;
		\item \textbf{Group B} and test separately on \textbf{Spatial} and \textbf{Cognitive};
	\end{itemize}
	\item As a baseline, we compared the above results to the earlier experiment shown in~\cref{fig:samnet_cog_detailed}, which can be viewed as training jointly on \textbf{All} and testing on each group at the leaf level separately (All-task training).
\end{itemize}

The results of these experiments are shown in~\cref{fig:cog_reasoning_transfer}.

@@

First, notice that for each of the \textbf{Basic} and \textbf{Obj-Attr} task families, the accuracy is near-perfect in all cases, suggesting that each contains the most primitive tasks and therefore do not benefit from training with other task families.
With \textbf{Spatial}, we see a small improvement showing that there is some benefit due to joint training with other task families.
Two groups that demonstrated a huge improvement of more than 25 points are \textbf{Compare} and \textbf{Cognitive}.
The former saw an accuracy jump from 58.6\% by training on samples from that family alone
to 97.0\% when training on all samples. To further emphasize this behavior, notice that
just joining \textbf{Compare} with \textbf{Obj-Attr} and \textbf{Basic} already causes a significant accuracy jump to 92.3\%.
In hindsight, this is not surprising, as the questions in \textbf{Compare} are composed
of fragments of questions given by \textbf{Basic} and \textbf{Obj-Attr}, and therefore can leverage the reasoning strategies developed there to reason about questions in \textbf{Compare}.
Lastly, for the \textbf{Spatial} family, we again see the benefits of joint training with all questions (68.6\% to 95.7\%) but in this case there
is a slight loss incurred by including everything. As seen in the figure, just jointly training with \textbf{Spatial} alone is sufficient to get a boost in accuracy (97.1\%). To summarize, while joint training helps, one needs to determine how much of correlation is present with the other tasks.

>>>>

\subsection{Reasoning transfer in CLEVR-CoGenT}
\label{sec:reasoning-transfer-clevr}
In the following experiments we have relied on the CoGent-A variant only.
Using the question groups (categories) defined by the authors, we reorganized training and testing splits with the goal of measuring whether mastering reasoning for one group can help with learning of the other.

@@

Next, we trained and tested 5 SAMNet instances, each on a single task group $t$ ("Train on t only").
Interestingly, we observe significant accuracy drops (from 18 up to almost 35 points) in all task groups aside of the \textit{Query Attribute}.
We hypotesize this results from the fact that mastering visual grounding and recognizing object attributes (being the core of the latter task) is a necessary skill the system needed by the other tasks.
Thus training on \textit{Query Attribute} helps others, whereas without any training on that particular group the system struggles with developing such a skill.

@@

\subsection{Reasoning transfer in COG Canonical}
\label{sec:reasoning-transfer-cog}
In order to further investigate whether reasoning transfer was effective in leveraging information gained by training a task family at a higher level of the hierarchy we also conducted a setof  experiments using the Canonical variant of COG.
Order of the performed experiments is analogical to the one from previous section.
First column (\cref{fig:cog_reasoning_transfer}, "Train on All") is our baseline, i.e.
we train a single SAMNet instance on all tasks and test on each of the five task groups from the task hierarchy presented in \cref{fig:task-groups}.
Please note that those results are in fact weighted averages when aggregating accuracies of our model on the Canonical variant (\cref{fig:samnet_cog_detailed}).

@@

Next, we experimented with the "Train on t only" setup, i.e. training and testing each of 5 SAMNet instances on a single task group $t$ (i.e. single leaf of the proposed hierarchy).
For the \textit{Basic}, \textit{Spatial} and \textit{Obj-Attr} task groups the accuracy is matching the baseline, suggesting that each group contain tasks sufficient for developing all necessary skills.
Therefore models trained on those will not benefit from any additional training on other task groups.
However, the accuracy drops in the models trained only on tasks from the \textit{Compary} and \textit{Cognitive} task groups suggests the oposite.

For that reason we have performed the third set of experiments ("Train on Group A/B"), where we trained two SAMNet instances: first on tasks from the \textit{Group A} aggregate, tested next on each task from the lowest groups (i.e. \textit{Basic}, \textit{Obj-Attr} and \textit{Compare}) separately; and second on tasks from \textit{Group B}, tested separately on \textit{Spatial} and \textit{Cognitive}.
As expected, training on other tasks from group \textit{Group A} enabled the model to significantly improve on \textit{Compare} task (from 58.6\% to 92.3\%).
Similar improvement (from 68.8\% to 97.1\%) was observed in the second model, when trained on \textit{Group B} and tested on \textit{Cognitive}.
Moreover, the achieved accuracy is in fact higher by 1.4 point from the one achieved when training on all tasks (95.7\%).
This suggest that the model might have been struggling with mastering all that one when faced all kind of questions, maybe lacking the capacity.
Finally, there is an interesting effect for \textit{Obj-Attr} and \textit{Spatial}: training on the whole group A/B seems not to help the models, and when comparing those accuracies with the ones achieved by the models trained on task $t$ only we observe some slight drops (around 0.3-0.5 point).
Understanding that effect seems to require further investigation on dependencies between the resoning required by the particular questions and reasoning transfer.

%With \textbf{Spatial}, we see a small improvement showing that there is some benefit due to joint training with other task families.
%To further emphasize this behavior, notice that just joining \textbf{Compare} with \textbf{Obj-Attr} and \textbf{Basic} already causes a significant accuracy jump to 92.3\%.
%In hindsight, this is not surprising, as the questions in \textbf{Compare} are composed of fragments of questions given by \textbf{Basic} and \textbf{Obj-Attr}, and therefore can leverage the reasoning strategies developed there to reason about questions in \textbf{Compare}.
% Lastly, for the \textbf{Spatial} family, we again see the benefits of joint training with all questions (68.6\% to 95.7\%) but in this case there is a slight loss incurred by including everything. As seen in the figure, just jointly training with \textbf{Spatial} alone is sufficient to get a boost in accuracy (97.1\%). To summarize, while joint training helps, one needs to determine how much of correlation is present with the other tasks.
